---
title: "Reliability report in one repetition maximum test"
author: "Aleksander, Vemund, Pernille and Tonje"
date: "2021"
output: html_document
---

## Introduction 
When testing in exercise physiology you need tests that are reliable. This means that you need to produce the same results each time you are doing a test. To be able to do so you need to standardize the test protocol and make sure that the participants have done the same things before pre and post testing. The elimination of learning progression, fatigue etc are therefore important to be able to produce reliable results. To get clean results we often look at the subject variation from pre to post.A good measure of reliability is the typical error, a percentage of the mean. This means the mean variation for the individuals standard deviation from pre to post test (Hopkins, 2000). 

## Method 
Nine healthy sports students participated in the reliability test. They performed a one repetition maximum strength test in leg press two times. These trials were performed one week apart, the same time of the day. Activity the day before and energy intake at test day were individually standardized. 1RM leg press test was performed with the Keiser Air 300 leg press machine. To establish reliability from pre to post test participants had the same test leader. Also the amount and gender of observers in the room was the same for both test days (Halperin et al., 2015). 

#### Preperation
Before 1RM test the Keiser Air 300 machine were individually customized. The participants performed the test without shoes with separate leg plates and 90 degrees in the knee joint. Seat length and leg position on the footplate were noted for each individual to ensure reproducibility from pre to post test. 

Three warm-up sets were performed before the 1RM test. It consisted of 8, 6 and 4 repetitions with increasing resistance and one minute rest in between. The resistance of the warm-up sets were noted. 

#### One repetition maximum test (1RM)
After the warm-up sets the participants rested for one minute. The participants had a maximum of five trials to reach 1RM. In between each trial they rested for three minutes. The resistance was determined by both the participant and test leader. Resistance of each trial was noted. 


## Results and discussion

```{r, message=FALSE, warning=FALSE, echo=FALSE}

library(readxl)
library (tidyverse)
library (knitr)
library (kableExtra)

  reliability_data <- read_excel("reliability_data.xlsx") %>%
  mutate (diff = Pre - Post) %>%
  
  summarise (s = round(sd (diff), 2),
    te = round(s / sqrt (2), 2), 
    m = round(mean (c(Pre, Post)), 2), 
    cv = round((te/m) * 100, 2))
  
  cv <- round(reliability_data$cv, 2)


```
```{r, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
reliability_data %>% 
  kable(format="html",  
        col.names = c("SD", "TE", "Mean", "CV"),
        caption = "Table 1. Descriptive statistics from 1RM pre and post test (n = 9).") %>%
        add_footnote(label= "SD = Standard deviation, TE = Typical error, CV = Coeffisient of variation",
                     notation = "none") %>%
  kable_styling(latex_options = "striped") 
```

The typical percentage error (CV) of the test was `r cv`%. This means that it is 2,14 % variation between each individual from pre to post test. This is a small variation and the test seems to be reliable. It may be explained by elimination of measurement errors because of  standardized preparation and test procedures. The small variation that was found in this reliability test may be explained by day-to-day variations which can be physiological and psychological factors, for example motivation and fatigue (Halperin et al., 2015) 


## References 
Hopkins W. G. (2000). Measures of reliability in sports medicine and science. Sports medicine (Auckland, N.Z.), 30(1), 1–15. https://doi.org/10.2165/00007256-200030010-00001

Halperin, I., Pyne, D. B., & Martin, D. T. (2015). Threats to internal validity in exercise science: a review of overlooked confounding variables. International journal of sports physiology and performance, 10(7), 823–829. https://doi.org/10.1123/ijspp.2014-0566 